---
title: "documentation for the transform function"
author: "31326"
date: "2023-10-19"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Description

This is documentation for demonstrating the usage and function of the transform_data in the repo

first we will look at our target data set iris

```{r, include=FALSE}
library(tidyverse)
library(ggplot2)
library(tidyverse)
library(knitr)
```

```{r dataset}
df <- iris
kable(head(iris))
```

we will the check the distribution of the iris data set, we found that there are not normalised

```{r}
hist(df$Sepal.Length)
hist(df$Sepal.Width)
hist(df$Petal.Length)
hist(df$Petal.Width)
```

we would transform the data by the transform_data function, with take a data and a method argument, indicating the dataset and method. we can use Z score or log to transform the function

```{r}
# transform_data.R
# this function is created for the iris dataset build-in within R, it normalize the 4 meaures of the iris dataset for potentiall categorisation tasks
# we are giving this function with a method argument to choose the type of normalisation to use
transform_data <- function(data, method = "scale") { 
  # Perform scale transformation
  if (method == "scale") {
    # existing scaling
    data$Sepal.Length <- scale(data$Sepal.Length) # normalise Sepal.Length with scale
    data$Sepal.Width <- scale(data$Sepal.Width) # normalise Sepal.Length with scale
    data$Petal.Length <- scale(data$Petal.Length) # normalise Sepal.Length with scale
    data$Petal.Width <- scale(data$Petal.Width) # normalise Sepal.Length with scale
  } else if (method == "log") {
    # logarithmic transformation
    data$Sepal.Length <- log(data$Sepal.Length) # normalise Sepal.Length with log
    data$Sepal.Width <- log(data$Sepal.Width) # normalise Sepal.Length with log
    data$Petal.Length <- log(data$Petal.Length) # normalise Sepal.Length with log
    data$Petal.Width <- log(data$Petal.Width) # normalise Sepal.Length with log
  }

  return(data)
}
```

we will use our function on the iris dataset, we see that the varibales has been centred

```{r}
df_trans <- transform_data(iris, method = "scale")
hist(df_trans$Sepal.Length)
hist(df_trans$Sepal.Width)
hist(df_trans$Petal.Length)
hist(df_trans$Petal.Width)
```